Modalities for Location Virtualization

### 2022-09-19 background.tex
../..

\subsection{Modal Logic}
\label{sec:backgroundonmodallogic}
The problem of needing to keep track of things being true in some contexts and not in others is hardly unique to virtual memory management, and is the general insight behind most flavors of modal logic.
\mytodo{Colin finish}

>>>>

\subsection{Modal Logic}
\label{sec:backgroundonmodallogic}
The problem of needing to keep track of things being true in some contexts and not in others is hardly unique to virtual memory management, and is the general insight behind most flavors of modal logic.

Most modal logics use unary operators on propositions to express that a logical claim $P$ is \emph{contingently} true in certain other circumstances, such as in other times~\cite{pnueli1977} or places~\cite{gordon2019modal}. That is, for different modal operators $M$, $M(P)$ may mean that $P$ is true in another time or place (which may be a physical place like another network node~\cite{murphy,gordon2019modal} or a conceptual place like a person's mind~\cite{hintikka1962knowledge}). A unifying concept across any modality is that they behave as applicative functors, typically satisfying (directly, or as a derived law):
\[ (P\rightarrow Q) \rightarrow M(P) \rightarrow M(Q)\]
\mytodo{Ismail, I think this means we have pure intro, $\square P \mathrel{-\ast} [r](\square P)$ if I'm using the right symbol for pure assertions }
Many modalities, so-called \emph{normal} modalities also possess introduction rules of the form $P\rightarrow M(P)$, the classic example being that if $P$ is true, then $P$ is \emph{necessarily} true ($\square P$)

Of particular interest for reasoning about virtual memory are modalities that permit \emph{naming} the alternate circumstances, prominently \emph{hybrid} modal logics~\cite{blackburn1995hybrid,areces2001hybrid}, which come equipped with assertions of the form $@_\ell(P)$ indicating that $P$ is true in the specific alternate circumstance (Kripke world) named by the \emph{nominal} $\ell$.

For our purposes, these are natural candidates to adapt for virtual memory management. We can reinterpret the notion of naming an alternate world slightly more loosely, and instead name \emph{address spaces} by the physical address of the page table root, since these structures are the physical representations of page tables. Thus in this paper we develop the notion that we can represent contingent truth of an assertion via $[r](P)$ indicating that $P$ holds in the address space rooted at physical address $r$.
The typical hybrid modality introduction rule, that $P$ and $\ell$ (indicating the current world is $\ell$) imply $@_\ell(P)$, has a natural analogue: knowing $P$ and that the current address space is $r$ (i.e., that $r$ is the current \texttt{cr3} value) suggests a way to construct $[r](P)$.

A relatively under-explored space of modal logics is the interaction of modal and substructural logics, in particular hybrid-style modalities in substructural logics, which has seen only minimal exploration and no prior application. We develop our ideas in the Iris program logic~\cite{jung2018}, as a result exploring additional subtleties that arise where the modality itself may entail ownership of resources, as well as interactions between our hybrid-style modality and substructural rules.  For example, Iris contains a number of modalities that distribute over separating conjunction, or for which resources can freely move into the modality (e.g., $\blacktriangleright(P)\ast Q \mathrel{-\ast} \blacktriangleright(P\ast Q)$). In our setting some of these rules apply while others do not. For example, in our setting, an assertion that involves no modalities is interpreted as holding in the current (active-on-the-CPU) address space, so clearly cannot move into arbitrary other address spaces --- unless guarded by another address space modality.

### 2022-09-19 syntax.tex
+
../..

\section{Syntax}
\label{sec:syntax}
\subsection{Registers, Memory Indices}
Programs we demonstrate in this paper requires accessing two types of computer resource: registers and memory. A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers,$\regset$. We use these identifiers to access the register values, $\regval \in \regvaltype$. Unlike registers, we do not abstract the memory indices as a special type but instead, for the sake of clarity and ease of representation, we show differently masked machine words, $\loc \in \Loc$, with the subscripts showing the length of a 64-bit machine word after masking, e.g. $\kw{w}_{12}$ is a 12-bit resource which can be obtained after masking 52-bit of a 64-bit word.
This simple syntax in \ref{fig:syntax} includes a simple syntax for values because our language is purely \textit{imperative}, i.e. an instruction does not return a value. Intentionally, we restrict any stream of instruction, $\instrs$, to eventually reduce to $\iskip$ instruction which is injected into unit. Our syntax capture the no-op ``$\iskip$'' and the sequencing construct ``$\iseq{\instr}{\instrs}$'' in their standard forms.
\input{fig-syntax}
\subsection{State}
\label{state}

>>>>

\section{Syntax}
\label{sec:syntax}
To develop our core logical ideas and explain their intepretation, we need only a handful of instructions, and a machine model corresponding to execution of x86-64 assembly instructions with virtual memory enabled on the CPU.

\subsection{Registers and Memory}
Programs we demonstrate in this paper requires accessing two types of computer resource: registers and memory. A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers,$\regset$. We use these identifiers to access the register values, $\regval \in \regvaltype$. Unlike registers, we do not abstract the memory indices as a special type but instead, for the sake of clarity and ease of representation, we show differently masked machine words, $\loc \in \Loc$, with the subscripts showing the length of a 64-bit machine word after masking, e.g. $\kw{w}_{12}$ is a 12-bit resource which can be obtained after masking 52-bit of a 64-bit word.
This simple syntax in \ref{fig:syntax} includes a simple syntax for values because our language is purely \textit{imperative}, i.e. an instruction does not return a value. Intentionally, we restrict any stream of instruction, $\instrs$, to eventually reduce to $\iskip$ instruction which is injected into unit. Our syntax capture the no-op ``$\iskip$'' and the sequencing construct ``$\iseq{\instr}{\instrs}$'' in their standard forms.
\input{fig-syntax}
